---
title: "**CBA stock return exploration**"
author: "Jackson Zhou"
output:
    html_document:
        theme: united
        highlight: kate
---

<style>
body {text-align: justify}
</style>

## Introduction

In this report, I will perform exploratory data analysis on daily Commonwealth Bank of Australia (CBA) stock return return prices from 03/01/2012 to 11/9/2020, as outlined in the assignment specifications.
This will involve:

- Plotting the prices and returns over time
- Calculating summary statistics for the price and return time series

## Importing packages

I will import any packages which might be useful for the analysis.

```{r Importing packages, message=FALSE, warning=FALSE}
library(tidyverse)  # Data manipulation
library(readxl)     # Load Excel files
library(data.table) # Shift vectors
library(e1071)      # Skewness and kurtosis functions
```

## Loading in and cleaning the data

To start, I will load in the CBA stock price data from the relevant cell range in the provided Excel file.
After this, I will generate the the time series for the returns, as specified by the following formula:
$$y_t=\log\frac{P_t}{P_{t-1}}$$

```{r Loading in and cleaning the data}
stock.data <- read_excel("Price History_Commonwealth bank.xlsx", sheet = "Sheet 1", range = "A12:B2210") %>% 
  rename("date" = "Exchange Date",
         "price" = "Close") %>% 
  mutate(return = round(log(price/shift(price, n = 1)), 5))
```

## Plotting the data

### Plotting the stock prices

Luckily, the `date` column has the correct data type, so I won't need to do any further work to get the plotting to work correctly.

```{r Plotting the stock prices, fig.align="center", fig.width=10}
stock.data %>% 
  ggplot(mapping = aes(x = date, y = price)) +
  geom_line(colour = "#303030") +
  labs(title = "CBA stock prices over time",
       x = "Date",
       y = "CBA stock price") +
  theme(plot.title = element_text(face = "bold", hjust = 0.5),
        panel.border = element_rect(colour = "black", fill = NA, size = 1))
```

Visualising the distribution of the stock prices using a histogram:

```{r Distribution of stock prices, fig.align="center"}
stock.data %>% 
  ggplot(mapping = aes(x = price)) +
  geom_histogram(bins = 30, colour = "black", fill = "#969696") +
  labs(title = "Distribution of CBA stock prices",
       x = "Stock price",
       y = "Frequency") +
  theme(plot.title = element_text(face = "bold", hjust = 0.5),
        panel.border = element_rect(colour = "black", fill = NA, size = 1))
```


### Plotting the stock returns

Using similar code from the previous section:

```{r Plotting the stock returns, fig.align="center", fig.width=10, warning=FALSE}
stock.data %>% 
  ggplot(mapping = aes(x = date, y = return)) +
  geom_line(colour = "#303030") +
  labs(title = "CBA stock returns over time",
       x = "Date",
       y = "CBA stock return") +
  theme(plot.title = element_text(face = "bold", hjust = 0.5),
        panel.border = element_rect(colour = "black", fill = NA, size = 1))
```

Visualising the distribution of the stock returns using a histogram:

```{r Distribution of stock returns, fig.align="center", warning=FALSE}
stock.data %>% 
  ggplot(mapping = aes(x = return)) +
  geom_histogram(bins = 30, colour = "black", fill = "#969696") +
  labs(title = "Distribution of CBA stock returns",
       x = "Stock return",
       y = "Frequency") +
  theme(plot.title = element_text(face = "bold", hjust = 0.5),
        panel.border = element_rect(colour = "black", fill = NA, size = 1))
```

## Calculating the summary statistics

Calculating basic summary statistics for the stock prices and returns:

```{r Basic summary statistics}
stock.data %>% 
  select(all_of(c("price", "return"))) %>% 
  summary()
```

Calculating skewness and kurtosis for the stock prices and returns:

```{r Skewness and kurtosis}
cat("For the CBA stock prices, the skewness is ", 
    round(skewness(stock.data$price), 5), 
    " and the kurtosis is ",
    round(kurtosis(stock.data$price), 5), 
    "\nFor the CBA stock returns, the skewness is ", 
    round(skewness(stock.data$return, na.rm = TRUE), 5), 
    " and the kurtosis is ",
    round(kurtosis(stock.data$return, na.rm = TRUE), 5), sep = "")
```
